---
title: "Temporal Chapter"
author: "Sam James"
date: "March 17, 2019"
output: html_document
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Temporal Chapter

The following is a collection of the scripts run to generate and analyse the data for temporal analyses of the juvenile sockeye diet data collected in 2015 and 2016.

Research Question: How do juvenile sockeye diets vary across the migration period and between years in two different coastal environments?

Sub-questions: 
1) Does the low foraging success observed in Chapter 1 (June 2015) persist throughout the migratory period and between years?
2) Does their diet show seasonal variation in species composition?
3) Does peak salmon outmigration overlap with peak zooplankton abundance?
4) (maybe, but too many questions!) Do the same patterns in prey selectivity observed in Chapter 1 persist through time?


## 1) Does the low foraging success observed in Chapter 1 (June 2015) persist throughout the migratory period and between years?

###Preparing the Data

```{r GFI, warning = FALSE}
library(tidyverse)
library(lubridate)
library(here)
library(visreg)

temporal <- read_csv(here("raw", "Temporal.csv"))

temporal$corrected.ww <- as.numeric(temporal$corrected.ww)
temporal$fishweight <- as.numeric(temporal$fishweight)
temporal$date <-as.Date(temporal$date)
temporal$year <- as.factor(temporal$year)

# Creating the date as a factor with specified levels
md.order <- c("May-14", "May-16", "May-21", "May-26", "May-27","Jun-03", "Jun-05", "Jun-06", "Jun-09", "Jun-11", "Jun-13", "Jun-14", "Jun-16", "Jun-20", "Jun-29")
temporal$monthday <- factor(format(temporal$date, "%b-%d"), levels = md.order)

#data wrangling
GFI <- temporal %>%
        select(ufn, location, site, date, year, monthday, fishweight, corrected.ww) %>%
        filter(!is.na(corrected.ww) & !is.na(fishweight)) %>%
        group_by(location, date) %>%
        mutate(gfi = (((corrected.ww)/1000)/fishweight)*100) %>%
        distinct()

# Plotting, separating data by location and year, with grey box for 'typical' GFI values
GFI.plot<- ggplot(data = GFI, aes(x = monthday, y = gfi)) +
        geom_boxplot(width=2, aes(fill=(location))) +
        theme_classic() +
        annotate("rect", xmin = 0, xmax = 16, ymin = 0.5, ymax = 1.5, alpha = .2) +
        facet_grid(location~year) +
        theme(
                axis.text.x = element_text(size = 10, angle = 90),
                axis.text.y = element_text(size = 10),
                axis.title.x = element_blank(),
                axis.title.y = element_text(size = 14),
                panel.background = element_rect(fill = "white", colour = "grey50"),
                panel.spacing.x = unit(2, "lines"),
                strip.background = element_blank(),
                strip.text = element_text(size = 16),
                legend.position = "none"
                ) +
        labs(y = "Gut Fullness Index", x = "Date") 

GFI.plot
```

### Testing for significance

I first looked to see if data were normally distributed: they are positively skewed, therefore data were sqrt transformed.

The ANOVA results show that there was a significant difference in GFI between locations, and that for JS, GFI was significantly higher in 2016.

```{r GFI ANOVA, warning = FALSE}

#ANOVA on GFI per location and year
hist(GFI$gfi)
#positive skew, therefore sqrt transform
GFI$sqrt_gfi <- sqrt(GFI$gfi)
hist(GFI$sqrt_gfi)


z <- lm(sqrt_gfi~location + year + location*year, data = GFI)
plot(z)
visreg(z)
anova(z)
#location and location*year interaction are significant meaning there is a significant difference
#between the two locaitons and the GFI was significantly higher in 2016 in JS only

#another way, same thing:
gfi.aov <- aov(sqrt_gfi ~ location + year + location*year, data = GFI)
summary(gfi.aov)


#to look at the effect of year at each location separately
gfi.DI <- GFI %>%
        filter(location == "DI")
gfi.JS <- GFI %>%
        filter(location == "JS")

#Anova for DI comparing years
z <- lm(sqrt_gfi~year, data = gfi.DI)
plot(z)
visreg(z)
anova(z)

#Anova for JS comparin years
z <- lm(sqrt_gfi~year, data = gfi.JS)
plot(z)
visreg(z)
anova(z)
```

******
## 2) Does their diet show seasonal variation in species composition?

This one is a bit more complicated. How do you best observe changes over time? I ran the Bray-Curtis analysis for all the data, and then separated by location. 

First we'll look at the data at the individual sample level, and then group them by sampling event later:

### Preparing the Data BY SAMPLE: Abundance, Biomass, and P/A

```{r Cluster: Importing Data, warning = FALSE}
library(tidyverse)
library(vegan)
library(lubridate)
library(dendextend)
library(here)


#Import files
temp <- read_csv(here("raw", "Temporal.csv"))
meta <- read_csv(here("processed", "temp_meta.csv"))
```

```{r Cluster: ABUNDANCE}
#Data wrangling:-------#

#using detailed level of taxonomy (plot.taxon.d), for coarser resolution use plot.taxon
#setting up the data and apply an sqrt arcsine tranformation to rel abund data
#Calculating abundance data at the individual sample level
#Removing sample "2015-05-16-D07-SO-27" because it's 100% dissimilar from all other samples (4 Thysanoessa spinifera and nothing else) and messes up NMDS
abund1 <- temp %>%
        select(semsp_id = semsp.no, ufn, date, sample.int, location, site, fullness.est, plot.taxon.d, count) %>%
        separate(date, c("year", "month", "day"), sep = "-") %>%
        unite(year_int, c(year, sample.int), remove = FALSE) %>%
        unite(year_month, c(year, month), remove = FALSE) %>%
        filter(semsp_id != "2015-05-16-D07-SO-27", count != "Na", !plot.taxon.d %in% c('NA', 'Digested', 'Anisakis'), fullness.est >5) %>% 
        group_by(semsp_id, year, month, year_month, year_int, sample.int, site, location, plot.taxon.d) %>%
        summarise(total = sum(count)) %>%
        mutate(relabund = (total/sum(total))) %>%
        mutate(arcsin = ((2/pi)*asin(sqrt(relabund)))) %>%
        select(semsp_id, year, month, year_month, year_int, sample.int, site, location, plot.taxon.d, arcsin)

#rearranging rows of taxonomic data into columns by plot taxon, making all NA's = 0, selecting only taxonomic columns
clusdata_abund<- spread(abund1, key = plot.taxon.d, value = arcsin) 
clusdata_abund[is.na(clusdata_abund)] <- 0

#converting it to a dataframe and setting rownames as the first column and then only selecting the taxonomic data columns
# as part of the data prep for Bray-Curtis
clusdata_abund <- data.frame(clusdata_abund)
rownames(clusdata_abund)=clusdata_abund[,1]    
bc_abund <- clusdata_abund[9:78]

#edit metadata to only contain same ID's as in the temporal fish data
metadata1 <- semi_join(meta, clusdata_abund, by = "semsp_id")
metadata <- merge(metadata1, clusdata_abund[,c("semsp_id", "year_month", "year_int", "sample.int")], by = "semsp_id") %>%
        select(1, 4, 2:3, 20:21, 5:7, 10:11, 8:9, 12:19)

```

We may also want to break it down by location to observe seasonal changes:

```{r: ABUNDANCE: DI vs JS}
#----DI-----
abund_DI <- temp %>%
        select(semsp_id = semsp.no, ufn, date, sample.int, location, site, fullness.est, plot.taxon.d, count) %>%
        separate(date, c("year", "month", "day"), sep = "-") %>%
        unite(year_int, c(year, sample.int), remove = FALSE) %>%
        unite(year_month, c(year, month), remove = FALSE) %>%
        filter(semsp_id != "2015-05-16-D07-SO-27", count != "Na", !plot.taxon.d %in% c('NA', 'Digested', 'Anisakis'), fullness.est >5, location == "DI") %>% 
        group_by(semsp_id, year, month, year_month, year_int, sample.int, plot.taxon.d) %>%
        summarise(total = sum(count)) %>%
        mutate(relabund = (total/sum(total))) %>%
        mutate(arcsin = ((2/pi)*asin(sqrt(relabund)))) %>%
        select(semsp_id, year, month, year_month, year_int, sample.int, plot.taxon.d, arcsin)

#rearranging rows of taxonomic data into columns by plot taxon, making all NA's = 0, selecting only taxonomic columns
clusdata_abund_DI<- spread(abund_DI, key = plot.taxon.d, value = arcsin) 
clusdata_abund_DI[is.na(clusdata_abund_DI)] <- 0

#converting it to a dataframe and setting rownames as the first column and then only selecting the taxonomic data columns
# as part of the data prep for Bray-Curtis
clusdata_abund_DI <- data.frame(clusdata_abund_DI)
rownames(clusdata_abund_DI)=clusdata_abund_DI[,1]    
bc_abund_DI <- clusdata_abund_DI[7:64]

#edit metadata to only contain same ID's as in the fish data
metadata1 <- semi_join(meta, clusdata_abund_DI, by = "semsp_id")
metadata_DI <- merge(metadata1, clusdata_abund_DI[,c("semsp_id", "year_month", "year_int", "sample.int")], by = "semsp_id") %>%
        select(1, 4, 2:3, 20:21, 5:7, 10:11, 8:9, 12:19)

#-----JS-----
abund_JS <- temp %>%
        select(semsp_id = semsp.no, ufn, date, sample.int, location, site, fullness.est, plot.taxon.d, count) %>%
        separate(date, c("year", "month", "day"), sep = "-") %>%
        unite(year_int, c(year, sample.int), remove = FALSE) %>%
        unite(year_month, c(year, month), remove = FALSE) %>%
        filter(semsp_id != "2015-05-16-D07-SO-27", count != "Na", !plot.taxon.d %in% c('NA', 'Digested', 'Anisakis'), fullness.est >5, location == "JS") %>% 
        group_by(semsp_id, year, month, year_month, year_int, sample.int, plot.taxon.d) %>%
        summarise(total = sum(count)) %>%
        mutate(relabund = (total/sum(total))) %>%
        mutate(arcsin = ((2/pi)*asin(sqrt(relabund)))) %>%
        select(semsp_id, year, month, year_month, year_int, sample.int, plot.taxon.d, arcsin)

#rearranging rows of taxonomic data into columns by plot taxon, making all NA's = 0, selecting only taxonomic columns
clusdata_abund_JS<- spread(abund_JS, key = plot.taxon.d, value = arcsin) 
clusdata_abund_JS[is.na(clusdata_abund_JS)] <- 0

#converting it to a dataframe and setting rownames as the first column and then only selecting the taxonomic data columns
# as part of the data prep for Bray-Curtis
clusdata_abund_JS <- data.frame(clusdata_abund_JS)
rownames(clusdata_abund_JS)=clusdata_abund_JS[,1]    
bc_abund_JS <- clusdata_abund_JS[7:62]

#edit metadata to only contain same ID's as in the fish data
metadata1 <- semi_join(meta, clusdata_abund_JS, by = "semsp_id")
metadata_JS <- merge(metadata1, clusdata_abund_JS[,c("semsp_id", "year_month", "year_int", "sample.int")], by = "semsp_id") %>%
        select(1, 4, 2:3, 20:21, 5:7, 10:11, 8:9, 12:19)
```

Now for the biomass data:

```{r Biomass}
##BIOMASS DATA:

#Calculating biomass for each individual sample
#Again, Removing sample "2015-05-16-D07-SO-27" because it's 100% dissimilar from all other samples (4 Thysanoessa spinifera and nothing else) and messes up NMDS
biom1 <- temp %>%
        select(ufn, date, sample.int, location, site, semsp_id = semsp.no, region, fullness.est, plot.taxon.d, corrected.weight) %>%
        separate(date, c("year", "month", "day"), sep = "-") %>%
        unite(year_int, c(year, sample.int), remove = FALSE) %>%
        unite(year_month, c(year, month), remove = FALSE) %>%
        filter(semsp_id != '2015-05-16_D07_27', corrected.weight != "Na", !plot.taxon.d %in% c('NA', 'Digested', 'Anisakis'), fullness.est >5) %>% 
        group_by(semsp_id, year, month, year_month, year_int, location, plot.taxon.d) %>%
        mutate_at('corrected.weight', as.numeric) %>%
        summarise(total = sum(corrected.weight)) %>%
        mutate(relbiom = (total/sum(total))) %>%
        mutate(arcsin = ((2/pi)*asin(sqrt(relbiom)))) %>%
        select(semsp_id, year, month, year_month, year_int, location, plot.taxon.d, arcsin)

#rearranging rows of taxonomic data into columns by plot taxon, making all NA's = 0, selecting only taxonomic columns
clusdata_biom<- spread(biom1, key = plot.taxon.d, value = arcsin) 
clusdata_biom[is.na(clusdata_biom)] <- 0

#converting it to a dataframe and setting rownames as the first column and then only selecting the taxonomic data columns
# as part of the data prep for Bray-Curtis
clusdata_biom <- data.frame(clusdata_biom)
rownames(clusdata_biom)=clusdata_biom[,1]    
bc_biom <- clusdata_biom[7:76]
```

And again, breaking down the biomass data into separate dataframes for DI vs JS:

```{r Biomass: DI vs JS}
##BIOMASS DATA:

#---DI--------:

#Calculating biomass for each individual sample in DI
biom_DI <- temp %>%
        select(ufn, date, sample.int, location, semsp_id = semsp.no, region, fullness.est, plot.taxon.d, corrected.weight) %>%
        separate(date, c("year", "month", "day"), sep = "-") %>%
        unite(year_int, c(year, sample.int), remove = FALSE) %>%
        unite(year_month, c(year, month), remove = FALSE) %>%
        filter(semsp_id != '2015-05-16_D07_27', corrected.weight != "Na", !plot.taxon.d %in% c('NA', 'Digested', 'Anisakis'), fullness.est >5, location == "DI") %>% 
        group_by(semsp_id, year, month, year_month, year_int, sample.int, plot.taxon.d) %>%
        mutate_at('corrected.weight', as.numeric) %>%
        summarise(total = sum(corrected.weight)) %>%
        mutate(relbiom = (total/sum(total))) %>%
        mutate(arcsin = ((2/pi)*asin(sqrt(relbiom)))) %>%
        select(semsp_id, year, month, year_month, year_int, sample.int, plot.taxon.d, arcsin)

#rearranging rows of taxonomic data into columns by plot taxon, making all NA's = 0, selecting only taxonomic columns
clusdata_biom_DI <- spread(biom_DI, key = plot.taxon.d, value = arcsin) 
clusdata_biom_DI[is.na(clusdata_biom_DI)] <- 0

#converting it to a dataframe and setting rownames as the first column and then only selecting the taxonomic data columns
# as part of the data prep for Bray-Curtis
clusdata_biom_DI <- data.frame(clusdata_biom_DI)
rownames(clusdata_biom_DI)=clusdata_biom_DI[,1]    
bc_biom_DI <- clusdata_biom_DI[7:65]


#------JS-------:

#Calculating biomass for each individual sample in JS
biom_JS <- temp %>%
        select(ufn, date, sample.int, location, semsp_id = semsp.no, region, fullness.est, plot.taxon.d, corrected.weight) %>%
        separate(date, c("year", "month", "day"), sep = "-") %>%
        unite(year_int, c(year, sample.int), remove = FALSE) %>%
        unite(year_month, c(year, month), remove = FALSE) %>%
        filter(semsp_id != '2015-05-16_D07_27', corrected.weight != "Na", !plot.taxon.d %in% c('NA', 'Digested', 'Anisakis'), fullness.est >5, location == "JS") %>% 
        group_by(semsp_id, year, month, year_month, year_int, sample.int, plot.taxon.d) %>%
        mutate_at('corrected.weight', as.numeric) %>%
        summarise(total = sum(corrected.weight)) %>%
        mutate(relbiom = (total/sum(total))) %>%
        mutate(arcsin = ((2/pi)*asin(sqrt(relbiom)))) %>%
        select(semsp_id, year, month, year_month, year_int, sample.int, plot.taxon.d, arcsin)

#rearranging rows of taxonomic data into columns by plot taxon, making all NA's = 0, selecting only taxonomic columns
clusdata_biom_JS <- spread(biom_JS, key = plot.taxon.d, value = arcsin) 
clusdata_biom_JS[is.na(clusdata_biom_JS)] <- 0

#converting it to a dataframe and setting rownames as the first column and then only selecting the taxonomic data columns
# as part of the data prep for Bray-Curtis
clusdata_biom_JS <- data.frame(clusdata_biom_JS)
rownames(clusdata_biom_JS)=clusdata_biom_JS[,1]    
bc_biom_JS <- clusdata_biom_JS[7:62]
```


### Cluster Analyses - ALL DATA

Code to run the Bray-Curtis Analysis, then test for significant clusters using SIMPROF, then look for significance between clusters using ANOSIM and see which species contribute to significant differences using SIMPER. 

The code below is for all the data first, and then separated into DI and JS.

```{r Cluster: Bray-Curtis}
#----------Bray-Curtis----------#

#vegdist uses the vegan package to calculate the distances between variables
bc_dist <- vegdist(bc_abund, method = "bray")

#hclust creates the dendrogram by calculating the amount of dissimilarity between clusters,
# average calculates the distance between all the different values in each group and then averages them
bc_clust <- hclust(bc_dist, method = "average")


#------------Plotting-------------#

# plotting the Bray-Curtis dissimilarity, I made it as a dendrogram to flatten out the bottom
den_temp_clust = as.dendrogram(bc_clust)
plot(den_temp_clust)

#Colouring (from https://rstudio-pubs-static.s3.amazonaws.com/68544_06343669257d4f35aaca449f9ff1e6f7.html)
#creating a colour palette with dif colours for DI and JS
cbPalette <- c("tomato3", "steelblue")
colorCode <- c(DI=cbPalette[1], JS = cbPalette[2])

labels_colors(den_temp_clust) <- colorCode[clusdata_abund$location][order.dendrogram(den_temp_clust)]
par(cex = 0.7, mar=c(7,4,1,1))
plot(den_temp_clust, ylab = "Bray-Curtis Dissimilarity")

```

```{r Cluster: Simprof, results = 'hide'}
##--------SIMPROF TEST-----------##
##
library(clustsig)

#Calculates number of significant clusters using average linkage method
simprof_test <- simprof(bc_abund, num.expected=200, num.simulated=100,
                        method.cluster="average", method.distance="czekanowski", 
                        method.transform="identity", alpha=0.01, sample.orientation = "row", 
                        const =1 ,silent=FALSE, increment=10, undef.zero=T, 
                        warn.braycurtis = F)
simprof.plot(simprof_test)

```

```{r Cluster: ANOSIM/SIMPER}
##
##------ANOSIM/SIMPER-------##
##

#create the BC dissimilarity matrix
bc_dist <- vegdist(bc_abund, method = "bray")

attach(metadata)

#anosim using our dissimilarity matrix and our grouping variable: location
temp.ano <- anosim(bc_dist, location)
summary(temp.ano)
plot(temp.ano)

#anosim using our dissimilarity matrix and our grouping variable: year
temp.ano <- anosim(bc_dist, year)
summary(temp.ano)
plot(temp.ano)

#anosim using our dissimilarity matrix and our grouping variable: month
temp.ano <- anosim(bc_dist, month)
summary(temp.ano)
plot(temp.ano)

#anosim using our dissimilarity matrix and our grouping variable: sample.int (interval from 1-3, with 1 being end of May, 2 = early June, 3 = late June)
temp.ano <- anosim(bc_dist, sample.int)
summary(temp.ano)
plot(temp.ano)

#simper analysis using the BC dissimilarity matrix and grouping variable
sim <- with(metadata, simper(bc_abund, location))

#produces table with average contribution to dissimilary, avg abundances in groups a and b, and the cumulative sum of the contributions to dissimilarity
summary(sim)

#lists most influential species and their cumulative contributions
sim 

#to get the overall dissimilarity between the clusters
lapply(sim, FUN = function(x){x$overall})
```

### Cluster Analyses - Discovery Islands

Same code as above to run the Bray-Curtis Analysis, then test for significant clusters using SIMPROF, then look for significance between clusters using ANOSIM and see which species contribute to significant differences using SIMPER. Looking at DI data only.

```{r DI Cluster: Bray-Curtis}
#----------Bray-Curtis----------#

#vegdist uses the vegan package to calculate the distances between variables
bc_dist <- vegdist(bc_abund_DI, method = "bray")

#hclust creates the dendrogram by calculating the amount of dissimilarity between clusters,
# average calculates the distance between all the different values in each group and then averages them
bc_clust <- hclust(bc_dist, method = "average")


#------------Plotting-------------#

# plotting the Bray-Curtis dissimilarity, I made it as a dendrogram to flatten out the bottom
den_temp_clust = as.dendrogram(bc_clust)
plot(den_temp_clust)

#Colouring (from https://rstudio-pubs-static.s3.amazonaws.com/68544_06343669257d4f35aaca449f9ff1e6f7.html)
#creating a colour palette with dif colours for DI and JS
cbPalette <- c("tomato3", "steelblue")
colorCode <- c(DI=cbPalette[1], JS = cbPalette[2])

labels_colors(den_temp_clust) <- colorCode[clusdata_abund_DI$location][order.dendrogram(den_temp_clust)]
par(cex = 0.7, mar=c(7,4,1,1))
plot(den_temp_clust, ylab = "Bray-Curtis Dissimilarity")

#creating a colour palette with dif colours for time intervals 1-3 (end of may, early to mid June, late june)
cbPalette.int <- c("chartreuse3", "deepskyblue", "dodgerblue4", "gold2", "darkorange3", "brown1")
colorCode.int <- c("2015_1"=cbPalette.int[1], "2015_2"=cbPalette.int[2], "2015_3"=cbPalette.int[3], "2016_1"=cbPalette.int[4], "2016_2"=cbPalette.int[5], "2016_3"=cbPalette.int[6])

# assigning the colorcode to the spatial cluster data and ordering it so that the x-axis labels are coloured according to region, requires package dendextend
labels_colors(den_temp_clust) <- colorCode.int[clusdata_abund_DI$year_int][order.dendrogram(den_temp_clust)]
par(cex = 0.7, mar=c(7,4,1,1))
plot(den_temp_clust, ylab = "Bray-Curtis Dissimilarity")

```

```{r DI Cluster: Simprof, results = 'hide'}
##--------SIMPROF TEST-----------##
##
library(clustsig)

#Calculates number of significant clusters using average linkage method
simprof_test <- simprof(bc_abund_DI, num.expected=200, num.simulated=100,
                        method.cluster="average", method.distance="czekanowski", 
                        method.transform="identity", alpha=0.01, sample.orientation = "row", 
                        const =1 ,silent=FALSE, increment=10, undef.zero=T, 
                        warn.braycurtis = F)
simprof.plot(simprof_test)

```

```{r DI Cluster: ANOSIM/SIMPER}
##
##------ANOSIM/SIMPER-------##
##

#create the BC dissimilarity matrix
bc_dist <- vegdist(bc_abund_DI, method = "bray")

attach(metadata_DI)

#anosim using our dissimilarity matrix and our grouping variable: year
temp.ano <- anosim(bc_dist, year)
summary(temp.ano)
plot(temp.ano)

#anosim using our dissimilarity matrix and our grouping variable: month
temp.ano <- anosim(bc_dist, month)
summary(temp.ano)
plot(temp.ano)

#anosim using our dissimilarity matrix and our grouping variable: sample.int (interval from 1-3, with 1 being end of May, 2 = early June, 3 = late June)
temp.ano <- anosim(bc_dist, sample.int)
summary(temp.ano)
plot(temp.ano)

#simper analysis using the BC dissimilarity matrix and grouping variable
sim <- with(metadata_DI, simper(bc_abund_DI, year))
summary(sim) #produces table with average contribution to dissimilary, avg abundances in groups a and b, and the cumulative sum of the contributions to dissimilarity
sim #lists most influential species and their cumulative contributions


#to get the overall dissimilarity between the clusters
lapply(sim, FUN = function(x){x$overall})
```

### Cluster Analyses - Johnstone Strait

Same code as above to run the Bray-Curtis Analysis, then test for significant clusters using SIMPROF, then look for significance between clusters using ANOSIM and see which species contribute to significant differences using SIMPER. Looking at DI data only.

```{r JS Cluster: Bray-Curtis}
#----------Bray-Curtis----------#

#vegdist uses the vegan package to calculate the distances between variables
bc_dist <- vegdist(bc_abund_JS, method = "bray")

#hclust creates the dendrogram by calculating the amount of dissimilarity between clusters,
# average calculates the distance between all the different values in each group and then averages them
bc_clust <- hclust(bc_dist, method = "average")


#------------Plotting-------------#

# plotting the Bray-Curtis dissimilarity, I made it as a dendrogram to flatten out the bottom
den_temp_clust = as.dendrogram(bc_clust)
plot(den_temp_clust)

#Colouring (from https://rstudio-pubs-static.s3.amazonaws.com/68544_06343669257d4f35aaca449f9ff1e6f7.html)
#creating a colour palette with dif colours for 2015 and 2016
cbPalette <- c("tomato3", "steelblue")
colorCode <- c("2015"=cbPalette[1], "2016" = cbPalette[2])

labels_colors(den_temp_clust) <- colorCode[clusdata_abund_JS$year][order.dendrogram(den_temp_clust)]
par(cex = 0.7, mar=c(7,4,1,1))
plot(den_temp_clust, ylab = "Bray-Curtis Dissimilarity")

#creating a colour palette with dif colours for time intervals 1-3 (end of may, early to mid June, late june)
cbPalette.int <- c("chartreuse3", "deepskyblue", "dodgerblue4", "gold2", "darkorange3", "brown1")
colorCode.int <- c("2015_1"=cbPalette.int[1], "2015_2"=cbPalette.int[2], "2015_3"=cbPalette.int[3], "2016_1"=cbPalette.int[4], "2016_2"=cbPalette.int[5], "2016_3"=cbPalette.int[6])

# assigning the colorcode to the spatial cluster data and ordering it so that the x-axis labels are coloured according to region, requires package dendextend
labels_colors(den_temp_clust) <- colorCode.int[clusdata_abund_JS$year_int][order.dendrogram(den_temp_clust)]
par(cex = 0.7, mar=c(7,4,1,1))
plot(den_temp_clust, ylab = "Bray-Curtis Dissimilarity")

```

```{r JS Cluster: Simprof, results = 'hide'}
##--------SIMPROF TEST-----------##
##
library(clustsig)

#Calculates number of significant clusters using average linkage method
simprof_test <- simprof(bc_abund_JS, num.expected=200, num.simulated=100,
                        method.cluster="average", method.distance="czekanowski", 
                        method.transform="identity", alpha=0.01, sample.orientation = "row", 
                        const =1 ,silent=FALSE, increment=10, undef.zero=T, 
                        warn.braycurtis = F)
simprof.plot(simprof_test)

```

```{r JS Cluster: ANOSIM/SIMPER}
##
##------ANOSIM/SIMPER-------##
##

#create the BC dissimilarity matrix
bc_dist <- vegdist(bc_abund_JS, method = "bray")

attach(metadata_JS)

#anosim using our dissimilarity matrix and our grouping variable: year
temp.ano <- anosim(bc_dist, year)
summary(temp.ano)
plot(temp.ano)

#anosim using our dissimilarity matrix and our grouping variable: month
temp.ano <- anosim(bc_dist, month)
summary(temp.ano)
plot(temp.ano)

#anosim using our dissimilarity matrix and our grouping variable: sample.int (interval from 1-3, with 1 being end of May, 2 = early June, 3 = late June)
temp.ano <- anosim(bc_dist, sample.int)
summary(temp.ano)
plot(temp.ano)

#simper analysis using the BC dissimilarity matrix and grouping variable
sim <- with(metadata_JS, simper(bc_abund_JS, year))
summary(sim) #produces table with average contribution to dissimilary, avg abundances in groups a and b, and the cumulative sum of the contributions to dissimilarity
sim #lists most influential species and their cumulative contributions


#to get the overall dissimilarity between the clusters
lapply(sim, FUN = function(x){x$overall})
```


### Ordination

Then conduct BIOENV model to find best correlated environmental vairables. Then run NMDS and ordinate with environmental parameters as vectors.


******
## 3) Does peak salmon outmigration overlap with peak zooplankton abundance?

Below, I am trying to get a sense of the timing of peak zooplankton abundance. I am using the surface tows collected by the JSP, however there are also the surface and full water column tows from the oceanography program that could be considered. 

Not only is the timing interesting, but also the changes of biomass in the different size classes.

First I look at changes in zooplankton composition over time, then changes in abundance, then biomass.

```{r Zooplankton Abundance}
zoopcomp <- read_csv(here("raw", "zoopcomp.csv"))
zoopww <- read_csv(here("raw", "zoopww.csv"))

##
##SPECIES COMP-----------------
##

#data wrangling zoop relative abundance data for species comp
temp.abundz <- zoopcomp %>%
  mutate(year = format(date, "%Y")) %>%
  select(sampleID, date, year, location, site, plot.taxon.g, totcount, abundance) %>%
  filter(site == "D07" | site =="J07", totcount != "NA") %>%
  mutate(site = factor(site, levels = c("D07", "J07"))) %>%      
  group_by(location,site, sampleID) %>%
  mutate(relativeabundance = (totcount/sum(totcount))*100) %>%
  filter(relativeabundance >= 5)

#messing around with the date format to have it plot nicely on the x axis
temp.abundz$year <- format(as.factor(temp.abundz$year))
temp.abundz$date_m <- format(as.Date(temp.abundz$date), "%b-%d")
temp.abundz$date_m1 <- as.Date(temp.abundz$date_m, format = "%b-%d")

##plot relative abundance for each taxon for the taxa that contribute >5%, fill with taxon
temp.zoop.relabund <- ggplot(data = temp.abundz, aes(x = date_m1, y = abundance)) + 
  geom_col(aes(fill = plot.taxon.g), position = "fill", width = 3) +
  theme_classic()+
  facet_grid(vars(year), vars(location), scales = "free") +
  scale_fill_manual(
          values = c("darkorange1", "mediumseagreen", "mediumpurple3","goldenrod1", "royalblue3", "grey40"),
          labels = c("Barnacle", "Calanoid Copepod", "Cladoceran", "Euphausiid", "Oikopleura", "Other")
  ) +
  theme(
    axis.text.x = element_text(size = 12, angle = 90),
    axis.text.y = element_text(size = 14),
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 14, margin = margin(r=10)),
    panel.background = element_rect(fill = "white", colour = "grey50"),
    strip.background = element_blank(),
    strip.text = element_text(size = 16),
    legend.title=element_blank(),
    legend.text = element_text(size = 12)
  ) +
  labs(y = "Relative Abundance")

temp.zoop.relabund +  scale_x_date(date_breaks ="1 week", date_labels = "%b-%d")


#plotting zoop abundances as a bar graph
temp.zoop.abund <- ggplot(data = temp.abundz, aes(x = date_m1, y = abundance)) + 
  geom_bar(stat = "identity", width = 2) +
  theme_classic() +
  facet_grid(vars(year), vars(location), scales = "free_x") +
  theme(
    axis.text.x = element_text(size = 12, angle = 90),
    axis.text.y = element_text(size = 14),
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 14, margin = margin(r = 10)),
    panel.background = element_rect(fill = "white", colour = "grey50"),
    strip.background = element_blank(),
    strip.text = element_text(size = 16),
    legend.title=element_blank(),
    legend.text = element_text(size = 14)
  ) +
  labs(y = "Abundance (individ/m3)") +
  scale_x_date(date_breaks ="1 week", date_labels = "%b-%d")

temp.zoop.abund

#For setting up the data for plotting abundance: summarizing abundance per sample
zoop_abund_sum <- temp.abundz %>%
        group_by(date, date_m1, year, location, sampleID) %>%
        summarise(totabund = sum(abundance))

#plotting as line graph
zoop_abund_sum_p <- ggplot(data = zoop_abund_sum, aes(x = date_m1, y = totabund)) + 
        geom_line() +
        geom_point() +
        annotate("rect", xmin = 0, xmax = 16, ymin = 0.5, ymax = 1.5, alpha = .2) +
        theme_classic()+
        facet_grid(vars(year), vars(location), scales = "free_x") +
        theme(
                axis.text.x = element_text(size = 12, angle = 90),
                axis.text.y = element_text(size = 14),
                axis.title.x = element_blank(),
                axis.title.y = element_text(size = 14, margin = margin(r = 10)),
                panel.background = element_rect(fill = "white", colour = "grey50"),
                strip.background = element_blank(),
                strip.text = element_text(size = 16),
                legend.title=element_blank(),
                legend.text = element_text(size = 14)
        ) +
        labs(y = "Abundance (individ/m3)") +
        scale_x_date(date_breaks ="1 week", date_labels = "%b-%d")

zoop_abund_sum_p
```


```{r Zoop Biomass}
##--------BIOMASS-------------

#data wrangling zoop relative abundance data for species comp
zoopww$biomass <- as.numeric(as.character(zoopww$biomass))
temp.zoop.biom <- zoopww %>%
  mutate(year = format(date, "%Y")) %>%
  select(sampleID, date, year, location, site, sieve, biomass) %>%
  filter(site == "D07" | site == "J07", biomass != "NA") %>%
  mutate(site = factor(site, levels = c("D07", "J07"))) 

temp.zoop.biom$year <- format(as.factor(temp.zoop.biom$year))
temp.zoop.biom$date_m <- format(as.Date(temp.zoop.biom$date), "%b-%d")
temp.zoop.biom$date_m1 <- as.Date(temp.zoop.biom$date_m, format = "%b-%d")

#plotting biomass as a bar graph
temp.zoop.biomp <- ggplot(data = temp.zoop.biom, aes(x = date_m1, y = biomass)) + 
  geom_bar(stat = "identity", width = 2) +
  theme_classic()+
  facet_grid(vars(year), vars(location)) +
  theme(
    axis.text.x = element_text(size = 12, angle = 90),
    axis.text.y = element_text(size = 14),
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 14, margin = margin(r = 10)),
    panel.background = element_rect(fill = "white", colour = "grey50"),
    strip.background = element_blank(),
    strip.text = element_text(size = 16),
    legend.title=element_blank(),
    legend.text = element_text(size = 14)
  ) +
  labs(y = "Biomass (mg/m3)") +
  scale_x_date(date_breaks ="1 week", date_labels = "%b-%d")

temp.zoop.biomp

#For setting up the data for a line graph: summarizing biomass per sample rather than per size class
zoop_biom_sum <- temp.zoop.biom %>%
        group_by(date, date_m1, year, location, sampleID) %>%
        mutate(totbiom = sum(biomass))

#plotting as line graph
zoop_biom_sum_p <- ggplot(data = zoop_biom_sum, aes(x = date_m1, y = totbiom)) + 
        theme_classic()+
        geom_line() +
        geom_point() +
        geom_rect(data = data.frame(xmin = decimal_date(as.Date(c("2015-05-20"))),
                              xmax = decimal_date(as.Date(c("2015-06-01"))),
                              ymin = -Inf,
                              ymax = Inf),
            aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
            fill = "grey", alpha = 0.5) +
        facet_grid(vars(year), vars(location), scales = "free_x") +
        theme(
                axis.text.x = element_text(size = 12, angle = 90),
                axis.text.y = element_text(size = 14),
                axis.title.x = element_blank(),
                axis.title.y = element_text(size = 14, margin = margin(r = 10)),
                panel.background = element_rect(fill = "white", colour = "grey50"),
                strip.background = element_blank(),
                strip.text = element_text(size = 16),
                legend.title=element_blank(),
                legend.text = element_text(size = 14)
        ) +
        labs(y = "Biomass (mg/m3)") +
        scale_x_date(date_breaks ="1 week", date_labels = "%b-%d")

zoop_biom_sum_p

z.biom_size_DI <- zoop_biom_sum %>%
        filter(location == "DI") %>%
        group_by(sampleID, sieve) %>%
        mutate(sieve_biom = sum(biomass))

z.biom_size_JS <- zoop_biom_sum %>%
        filter(location == "JS") %>%
        group_by(sampleID, sieve) %>%
        mutate(sieve_biom = sum(biomass))

#plotting as line graph based on sieve for DI
z.biom_size_DIp <- ggplot(data = z.biom_size_DI, aes(x = date_m1, y = biomass)) + 
        geom_line() +
        geom_point() +
        theme_classic()+
        facet_grid(vars(sieve), vars(year), scales = "free_x") +
        theme(
                axis.text.x = element_text(size = 12, angle = 90),
                axis.text.y = element_text(size = 14),
                axis.title.x = element_blank(),
                axis.title.y = element_text(size = 14, margin = margin(r = 10)),
                panel.background = element_rect(fill = "white", colour = "grey50"),
                strip.background = element_blank(),
                strip.text = element_text(size = 16),
                legend.title=element_blank(),
                legend.text = element_text(size = 14)
        ) +
        labs(y = "Biomass (mg/m3)") +
        scale_x_date(date_breaks ="1 week", date_labels = "%b-%d")

z.biom_size_DIp

#plotting as line graph based on sieve for JS
z.biom_size_JSp <- ggplot(data = z.biom_size_JS, aes(x = date_m1, y = sieve_biom)) + 
        geom_line() +
        geom_point() +
        theme_classic()+
        facet_grid(vars(sieve), vars(year), scales = "free_x") +
        theme(
                axis.text.x = element_text(size = 12, angle = 90),
                axis.text.y = element_text(size = 14),
                axis.title.x = element_blank(),
                axis.title.y = element_text(size = 14, margin = margin(r = 10)),
                panel.background = element_rect(fill = "white", colour = "grey50"),
                strip.background = element_blank(),
                strip.text = element_text(size = 16),
                legend.title=element_blank(),
                legend.text = element_text(size = 14)
        ) +
        labs(y = "Biomass (mg/m3)") +
        scale_x_date(date_breaks ="1 week", date_labels = "%b-%d")

z.biom_size_JSp
```





