---
title: "Temporal_Cluster"
author: "Sam James"
date: "March 17, 2019"
output: pdf_document
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Temporal Cluster Analysis of Juvenile Sockeye Diets

The following is a collection of the scripts run to generate and analyse the data for temporal cluster analyses of the juvenile sockeye diet data collected in 2015 and 2016.

## Abundance Data
###Preparing the Data

```{r Cluster: data wrangling, warning = FALSE}
library(tidyverse)
library(vegan)
library(lubridate)
library(dendextend)
library(here)


#Import files
temp <- read_csv(here("raw", "Temporal.csv"))
meta <- read_csv(here("processed", "temp_meta.csv"))


#Data wrangling:-------#

#using detailed level of taxonomy (plot.taxon.d), for coarser resolution use plot.taxon
#setting up the data and apply an sqrt arcsine tranformation to rel abund data
#Calculating abundance data at the individual sample level
temp_abund1 <- temp %>%
        select(semsp_id = semsp.no, ufn, date, sample.int, location, site, fullness.est, plot.taxon.d, count) %>%
        separate(date, c("year", "month", "day"), sep = "-") %>%
        unite(year_int, c(year, sample.int), remove = FALSE) %>%
        filter(count != "Na", !plot.taxon.d %in% c('NA', 'Digested', 'Anisakis'), fullness.est >5) %>% 
        group_by(semsp_id, ID, year, month, sample.int, site, location, plot.taxon.d) %>%
        summarise(total = sum(count)) %>%
        mutate(relabund = (total/sum(total))) %>%
        mutate(arcsin = ((2/pi)*asin(sqrt(relabund)))) %>%
        select(semsp_id, ID, year, month, sample.int, site, location, plot.taxon.d, arcsin)

#save
write.csv(temp_abund1, "./processed/temp_relabund_transformed.csv")

#rearranging rows of taxonomic data into columns by plot taxon, making all NA's = 0, selecting only taxonomic columns
temp_clusdata_abund<- spread(temp_abund1, key = plot.taxon.d, value = arcsin) 
temp_clusdata_abund[is.na(temp_clusdata_abund)] <- 0

#converting it to a dataframe and setting rownames as the first column and then only selecting the taxonomic data columns
# as part of the data prep for Bray-Curtis
temp_clusdata_abund <- data.frame(temp_clusdata_abund)
rownames(temp_clusdata_abund)=temp_clusdata_abund[,1]    
temp_bc_abund <- temp_clusdata_abund[8:77]

#edit metadata to only contain same ID's as in the temporal fish data
metadata1 <- semi_join(meta, temp_clusdata_abund, by = "semsp_id")
metadata <- merge(metadata1, temp_clusdata_abund[,c("semsp_id", "ID", "sample.int")], by = "semsp_id") %>%
        select(1, 4, 2:3, 20:21, 5:19)

```

### Cluster Analyses

Code to run the Bray-Curtis Analysis, then test for significant clusters using SIMPROF, then look for significance between clusters using ANOSIM and see which species contribute to significant differences using SIMPER. 

```{r Cluster: Bray-Curtis}
#----------Bray-Curtis----------#

#vegdist uses the vegan package to calculate the distances between variables
temp_bc_dist <- vegdist(temp_bc_abund, method = "bray")

#hclust creates the dendrogram by calculating the amount of dissimilarity between clusters,
# average calculates the distance between all the different values in each group and then averages them
temp_bc_clust <- hclust(temp_bc_dist, method = "average")


#------------Plotting-------------#

# plotting the Bray-Curtis dissimilarity, I made it as a dendrogram to flatten out the bottom
den_temp_clust = as.dendrogram(temp_bc_clust)
plot(den_temp_clust)

#Colouring (from https://rstudio-pubs-static.s3.amazonaws.com/68544_06343669257d4f35aaca449f9ff1e6f7.html)
#creating a colour palette with dif colours for DI and JS
cbPalette <- c("tomato3", "steelblue")
colorCode <- c(DI=cbPalette[1], JS = cbPalette[2])

labels_colors(den_temp_clust) <- colorCode[temp_clusdata_abund$location][order.dendrogram(den_temp_clust)]
par(cex = 0.7, mar=c(7,4,1,1))
plot(den_temp_clust, ylab = "Bray-Curtis Dissimilarity")

```

```{r Cluster: Simprof, results = 'hide'}
##--------SIMPROF TEST-----------##
##
library(clustsig)

#Calculates number of significant clusters using average linkage method
simprof_test <- simprof(temp_bc_abund, num.expected=200, num.simulated=100,
                        method.cluster="average", method.distance="czekanowski", 
                        method.transform="identity", alpha=0.01, sample.orientation = "row", 
                        const =1 ,silent=FALSE, increment=10, undef.zero=T, 
                        warn.braycurtis = F)
simprof.plot(simprof_test)

```

```{r Cluster: ANOSIM/SIMPER}
##
##------ANOSIM/SIMPER-------##
##

#create the BC dissimilarity matrix
temp_bc_dist <- vegdist(temp_bc_abund, method = "bray")

attach(metadata)

#anosim using our dissimilarity matrix and our grouping variable: location
temp.ano <- anosim(temp_bc_dist, location)
summary(temp.ano)
plot(temp.ano)

#anosim using our dissimilarity matrix and our grouping variable: year
temp.ano <- anosim(temp_bc_dist, year)
summary(temp.ano)
plot(temp.ano)

#anosim using our dissimilarity matrix and our grouping variable: month
temp.ano <- anosim(temp_bc_dist, month)
summary(temp.ano)
plot(temp.ano)

#anosim using our dissimilarity matrix and our grouping variable: sample.int (interval from 1-3, with 1 being end of May, 2 = early June, 3 = late June)
temp.ano <- anosim(temp_bc_dist, sample.int)
summary(temp.ano)
plot(temp.ano)

#simper analysis using the BC dissimilarity matrix and grouping variable
sim <- with(metadata, simper(temp_bc_abund, location))
summary(sim) #produces table with average contribution to dissimilary, avg abundances in groups a and b, and the cumulative sum of the contributions to dissimilarity
sim #lists most influential species and their cumulative contributions


#to get the overall dissimilarity between the clusters
lapply(sim, FUN = function(x){x$overall})
```



### Ordination

Then conduct BIOENV model to find best correlated environmental vairables. Then run NMDS and ordinate with environmental parameters as vectors.

